Reduced growth due to belowground sink limitation is not fully explained by reduced photosynthesis

### 2016-12-14 manuscript/reviewer_response_tree_physiology.Rmd

```{r, results='asis', echo=FALSE, warning=FALSE}
library(knitr)
opts_knit$set(root.dir = '../')
```

Dr Ram Oren
\
Editor-in-Chief
\
Tree Physiology

\
Dear Professor Mott,

We are grateful for the opportunity to resubmit our manuscript. We thank the reviewers for providing insightful comments to improve our manuscript. We have responded in detail to each comment, which has improved the overall quality of the manuscript. Below, we have itemized our response to each comment from both referees.

@@

### **Referee #1**

\
*Reviewer: 1. The role of nutrient defiency may have been underestimated...The relative contribution should be determined...*

\
*Response*: Unfortunately, we do not have the ability with this data set to determine the relative contribution of nutrient deficiency on the down regulation of photosynthesis. We were able to identify not one, but two processes that likely contributed to photosynthetic inhibition.  Both of these processes, nutrient limitation and starch accumulation, are discussed at length in the discussion. We absolutely agree that determining the contribution of each of these processes will provide valuable insights on how sink limitation impacts the carbon balance of plants.  By providing convincing empirical evidence of these processes, should give a framework for future work to hopefully quantify the contributions of each.

\
*Reviewer: 2. authors indicated that one of the principal question in the paper was to address which physiological process best explains the down-regulation of photosynthesis...the physiological processes involved have not been determined in the work*

\
*Response*: We agree with the reviewers point here. We have re-worded the first principle question to more accurately address what we were able to test with the presented data.

\
*Reviewer: 3. After reading the title, aim of the work and abstract, it seems that the manuscript focused specifically on carbon source-sink relations and regulation. In this regard, too much conclusions are derived from the presented data. Nevertheless, discussion and conclusions focused more accurately on the obtained results*

\
*Response*: We have clarified the role of sink limitation, via nutrient limitation and starch accumulation, on photosynthetic down regulation in the abstract. In addition to the response to comment #2, the focus of the beginning of the paper should now better align with the impacts of sink manipulation that are addressed at length in the discussion.

\
*Reviewer: 4. pp35: "to utilize in growth and storage”*

@@

\
*Response*: editorial change made

\
*Reviewer: 5. pp 40-45: Should be improved. Carbon losses are due to respiration and photorespiration.*

\
*Response*: We understand what the reviewer is suggesting but disagree with the suggested change.  Instead we have re-worded the sentence to say "C entering the planted through net photosynthesis must be balanced by C loss to respiration...".

\
*Reviewer: 6. pp 63: “…the sink limitation of growth has been explored by manipulations of C source and sink activity…. Examples of …that manipulate C source activity include… and partial defoliation..” (partial defoliation should impose a sink limitation to growth?) Compensatory increases of An are commonly found, but not allowing a higher C net gain leading to sink limitation.*

\
*Response*: Regardless of the downstream effects of defoliation on growth processes, the removal of leaves is first and foremost a source manipulation. This has been repeatedly shown and is a commonly utilized manipulation in fruit trees and crops species.

\
*Reviewer: 7. pp 365: No changes in stomatal conductance during the experiment in the described field conditions (Figure 2)?*

\
*Response*: Soil moisture for each seedling were maintained at field capacity, throughout the duration of the experiment, in order to minimize affects of reduced water availability on photosynthesis and growth. The consistent lack of water stress on theses seedlings (see Discussion "Changes in growth and physiology under sink limitation") was an successful implement of the experimental design.

\
*Reviewer: 8. pp 373: Modelled cumulative net leaf gain could be displayed in a table? (e.g. Table 1)*

\
*Response*: We have added modelled cumulative net leaf gain to Table 1. and refer to it in the Results section "Whole-plant C balance"

\
*Reviewer: 9. Table 1 and 2: Reorganize. Biomass, SLA, SRL and FRLD in one table, and chemistry parameters in a second table.*

\
*Response*: We agree with the reviewer.  Table 1 and 2 have been reorganized, as suggested, to match how the data are presented in the Results section.

\
*Reviewer: 10. Figure 4: Significant differences between 5 and 15 L soil volumes? Are SE or SD displayed?*

\
*Response*: Bar plots of Figure 4 are means and standard errors which are described in figure caption. Yes, the 5 l containers had smaller rates of A~sat~ than either 15 or 20 L containers. The 5 l containers also had the lowest measured rates overall, indicative of a strong impact of a small container size on sink strength.  For simplicity, we report the large overall difference in Free seedlings and seedlings in containers in terms of photosynthesis rates.

@@

### **Referee #2**

\
*Reviewer: 1.  It is not clear whether trees used in this experiment come from a nursery or if they were lifted from the field*

\
*Response*: We have clarified in "Experimental design" that seedlings were nursery seed grown in 40cm tall tube-stock.

@@

\
*Reviewer: 3 In relation to the “Photosynthetic parameters”, please, define Jmax and Vcmax.*

\
*Response*: We have defined both *V*~cmax~ and *J*~max~ in the section "Photosynthetic parameters"

\
*Reviewer: 4. In “Leaf, root and soil chemistry” indicate how plant samples were dried (temperature, drying time), how they were milled (type of mill used, model, etc.) and how soils were sieved (pore size) and milled (type of mill used, model, etc.)*

\
*Response*: We have clarified in “Leaf, root and soil chemistry” that samples were over dried at 60&degC to a constant mass and ball milled for analysis.We have also added the pore size, drying time and mill type to explanation of soil processing.

@@

\
*Reviewer: 5.  Clairify how did soil volume treatments relate to the 61 digitalized seedlings, which include different *Eucalyptus* species? Were the 61 digitized seedlings grown in different soil volumes*

\
*Response*: The digitized seedlings used here are a collection of digitized plants from differing environments including understory, glasshouse and plantation (see Duursma et al. 2012).  We selected 5 *Eucalpytus* species from this collection, which represent canopy structures of similar seedling types and sizes. Although the digitized plants are from separate experiments, they allow us to estimate self shading of the *Eucalyptus* seedlings in this experiment.  As this experiment had an ~4 month duration, leaf counts and total leaf area of seedlings were generally small. This resulted in only a modest adjustment of the self-shading factor to total canopy carbon uptake.

@@

\
*Reviewer: 6. Instead of using C concentration data (which they had), the authors assumed that C accounted for half of the final dry mass of seedlings. I think this is quite a strong assumption to make; differences in C concentration above 5% could be in place*

\
*Response*: Unfortunately, we do not have available C concentration data for plant tissues. Our plant chemistry data set includes leaf and root tissue nitrogen content and leaf TNC and 13C. Consequently, we have to assume a C concentration for plant tissues. Our use of a 50% C content is likely very close to any assumption of C content that we have to make. For example, an average value of C content of angiosperm woody tissues from the DRYAD wood C database is ~48%.

@@

\
*Reviewer: 7.  I would strongly recommend adopting a whole-plant approach and the consideration of all possible sources and sinks in mass-balance models. For example, root exudation can account for an important proportion of the total C budget of a plant, but yet this was not considered in the C budget calculations (nor even suggested as an important component to include in future models, see line 501)*

\
*Response*: The reviewer makes a very good point relating to how mass balance attempts should account for all possible fates of net C uptake. In this study, as in many, it is difficult to accurately measure every facet of C allocation throughout the plant and into the soil. As the reviewer point out, it is entirely possible that aspects of belowground sink limitation could have impacted patterns of root exudation in these seedlings. However, root exudation is just one of these possible fates, which also include woody tissue respiration, herbivory or VOC's. As we were not able measure all of these possible fates of net C uptake, we utilized the mass balance approach to estimate the residual difference between the modelled seedling C uptake and harvested C mass. It is for this reason that we report differences in CUE, as a more conservative approach in testing the coordination between photosynthesis and growth. Modelling the impacts of other potential fates of C, including root exudation, would require a priori knowledge of the effects of sink limitation/root restriction/container volume on each of these process. As so many of these processes are extremely difficult to measure, we do not yet have the ability to make such predictions. Possible impacts of root exudation are addressed in "Biomass partitioning under sink limitation", but we do not include it ubiquitously in the discussion because it is just one of many potential pathways of C loss.

@@

\
*Reviewer: 8.  I think future empirical and C models should evaluate the inter-relationship between the different sinks and sources AT THE WHOLE PLANT level, and not just the leaf level as indicated in lines 499-501.*

\
*Response*: We absolutely agree with the reviewer. We have made the editorial change to include the fate of C allocation at whole plant level, which includes the addition of root exudation from the previous comment.

\
*Reviewer: 9.  Lines 56-59. Yes, but beware the suggested approach also ignores part of the C budget of plants, because you are not accounting for root exudates.*

\
*Response*: see response to comment #7.  Our approach uses the residual C from mass balance to all other potential C losses.

### 2016-12-14 manuscript/manuscript_pve.Rmd
edits to manuscript and reviewer comments, made new Table 1 and 2, need to implement new tables in manuscript

Sink limitation is known to reduce plant growth, but it is not known how plant carbon (C) balance is affected, limiting our ability to predict growth under sink-limited conditions. We manipulated soil volume to impose sink limitation of growth in *Eucalyptus tereticornis* seedlings. Seedlings were grown in the field in containers of different sizes and planted flush to the soil alongside freely-rooted (Free) seedlings. Container volume negatively affected aboveground growth throughout the experiment, and light saturated rates of leaf photosynthesis were consistently lower in seedlings in containers (-26 %) compared to Free seedlings. Photosynthetic capacity was also significantly reduced in containerized seedlings, and was related to both reduced leaf nitrogen content and starch accumulation.  After 120 days, harvested biomass of Free seedlings was on average 84 % higher than seedlings in containers, but biomass distribution in leaves, stems and roots was not different. However, the reduction in net leaf photosynthesis over the growth period was insufficient to explain the reduction in growth, so that we also observed an apparent reduction in whole-plant C use efficiency (CUE) with container size. Our results show that sink limitation affects plant growth through feedbacks to both photosynthesis and CUE. Mass balance approaches to predicting plant growth under sink-limited conditions need to incorporate both of these feedbacks.

@@

\
Most current models of plant growth are based on the physical principle of mass balance: C entering the plant through photosynthesis must be balanced by C loss to respiration and C allocation to growth and storage. Plant growth is thus predicted as the outcome of the processes of photosynthesis, respiration and allocation. It is easy to see how source-limited growth can be implemented in such models: increasing source activity increases photosynthetic C uptake, flowing through to increased growth. It is less easy to know how to implement sink limitation of growth in such models, because it requires that we know whether, and how, sink limitation influences the contributing processes of photosynthesis, respiration and allocation. It has been suggested that sink limitation affects the C balance by causing a down-regulation of photosynthetic rate, possibly mediated by carbohydrate accumulation [@arp1991effects; @paul2001sink;@ainsworth2004testing; @ronchi2006growth; @nebauer2011photosynthesis], but there is also evidence that photosynthesis and growth can become strongly decoupled under sink limitation [@muller2011water;@leuzinger2013sink], suggesting that other processes must also be affected.

@@

\
Our goal in this paper was to impose a sink limitation of growth, by manipulating container size and thus rooting volume, and to examine how the C balance of the plants was affected. We grew *Eucalyptus tereticornis* seedlings in a range of container sizes in field conditions, using freely-rooted seedlings as a control for the container size treatments. Seedlings were maintained under well-watered conditions in order to isolate the effect of restricted soil volume from that of low water availability. We tracked leaf-level gas exchange, leaf carbohydrate accumulation and seedling allometry over the course of 4 months. We drove a simple whole-plant photosynthesis model with measured leaf gas exchange and interpolated leaf area data to estimated net canopy C uptake over the experimental period, and compared the result with total seedling biomass. We anticipated that both photosynthesis and growth would be reduced under belowground sink limitation. The principal questions that we addressed were: 1) which physiological process best explains the down-regulation of photosynthesis?, and 2) does this down-regulation fully explain the observed response of growth, or are other components of the C balance also affected?

@@

##Whole-plant C balance
Modelled cumulative net leaf C gain over 120 days varied two-fold across containers sizes (25.6 - 55.1 g C). The Free seedling control had five-fold greater modelled total C gain than the average of container treatments (213.6 - 42.7&pm;8.2 g C). For seedlings in containers, an average of 67 &pm;0.01 % of modelled C gain was not allocated to observed biomass, compared to 59.2 % for the Free seedling control. Consequently, the estimated CUE of containerized seedlings was lower than the Free seedling control (Figure `r cuefig`). As a result, the observed reductions in leaf *A*~n~ with decreasing soil volume when integrated across the 120 day experiment were not large enough to explain the reduction in observed seedling biomass across the container size treatments.

>>>>

Sink limitation is known to reduce plant growth, but it is not known how plant carbon (C) balance is affected, limiting our ability to predict growth under sink-limited conditions. We manipulated soil volume to impose sink limitation of growth in *Eucalyptus tereticornis* seedlings. Seedlings were grown in the field in containers of different sizes and planted flush to the soil alongside freely-rooted (Free) seedlings. Container volume negatively affected aboveground growth throughout the experiment, and light saturated rates of leaf photosynthesis were consistently lower in seedlings in containers (-26 %) compared to Free seedlings. Significant reductions in photosynthetic capacity in containerized seedlings was related to both reduced leaf nitrogen content and starch accumulation, indicating direct effects of sink limitation on photosynthetic down regulation. After 120 days, harvested biomass of Free seedlings was on average 84 % higher than seedlings in containers, but biomass distribution in leaves, stems and roots was not different. However, the reduction in net leaf photosynthesis over the growth period was insufficient to explain the reduction in growth, so that we also observed an apparent reduction in whole-plant C use efficiency (CUE) with container size. Our results show that sink limitation affects plant growth through feedbacks to both photosynthesis and CUE. Mass balance approaches to predicting plant growth under sink-limited conditions need to incorporate both of these feedbacks.

@@

\
Most current models of plant growth are based on the physical principle of mass balance: C entering the plant through net photosynthesis must be balanced by C loss to respiration and C allocation to growth and storage. Plant growth is thus predicted as the outcome of the processes of photosynthesis, respiration and allocation. It is easy to see how source-limited growth can be implemented in such models: increasing source activity increases photosynthetic C uptake, flowing through to increased growth. It is less easy to know how to implement sink limitation of growth in such models, because it requires that we know whether, and how, sink limitation influences the contributing processes of photosynthesis, respiration and allocation. It has been suggested that sink limitation affects the C balance by causing a down-regulation of photosynthetic rate, possibly mediated by carbohydrate accumulation [@arp1991effects; @paul2001sink;@ainsworth2004testing; @ronchi2006growth; @nebauer2011photosynthesis], but there is also evidence that photosynthesis and growth can become strongly decoupled under sink limitation [@muller2011water;@leuzinger2013sink], suggesting that other processes must also be affected.

@@

\
Our goal in this paper was to impose a sink limitation of growth, by manipulating container size and thus rooting volume, and to examine how the C balance of the plants was affected. We grew *Eucalyptus tereticornis* seedlings in a range of container sizes in field conditions, using freely-rooted seedlings as a control for the container size treatments. Seedlings were maintained under well-watered conditions in order to isolate the effect of restricted soil volume from that of low water availability. We tracked leaf-level gas exchange, leaf carbohydrate accumulation and seedling allometry over the course of 4 months. We drove a simple whole-plant photosynthesis model with measured leaf gas exchange and interpolated leaf area data to estimated net canopy C uptake over the experimental period, and compared the result with total seedling biomass. We anticipated that both photosynthesis and growth would be reduced under belowground sink limitation. The principal questions that we addressed were: 1) can we identify the main physiological processes that best explain the down-regulation of photosynthesis?, and 2) does this down-regulation fully explain the observed response of growth, or are other components of the C balance also affected?

@@

##Whole-plant C balance
Modelled cumulative net leaf C gain over 120 days varied three-fold across containers sizes (23.1 - 79.9 g C). The Free seedling control had five-fold greater modelled total C gain than the average of container treatments (213.6 - 42.7&pm;8.2 g C). For seedlings in containers, an average of 67 &pm;0.01 % of modelled C gain was not allocated to observed biomass, compared to 59.2 % for the Free seedling control. Consequently, the estimated CUE of containerized seedlings was lower than the Free seedling control (Figure `r cuefig`). As a result, the observed reductions in leaf *A*~n~ with decreasing soil volume when integrated across the 120 day experiment were not large enough to explain the reduction in observed seedling biomass across the container size treatments.
